Type Refinements for the Working Class

### 2015-11-20 paper.tex
Remove formalization of evaluation relation (#5)
Switch to informal meaning explanations for the LF (#5)

For any abt signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we can deploy
a generic logical framework. Recall from~\cite{sterling-morrison:2015} that
$\SCtx$ is the category of symbol-contexts and their injective, sort-preserving
renamings. Then, a \emph{judgment} shall be interpreted as an object in the presheaf
category $\Define{\mathbb{J}}{\Sets^\SCtx}$; in other words, a
judgment is identified with the intensional set of its renaming-sensitive
derivations. $\mathbb{J}$, like all presheaf categories, is a topos.

\begin{definition}[Hypothetical Judgment]
  For two judgments $\mathcal{J}_1$ and $\mathcal{J}_2$, we can form the
  hypothetical judgment $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$ using the
  exponential construction~\cite[p.~46]{maclane-moerdijk:1992}, which expresses
  the semantic consequence of $\mathcal{J}_2$ from $\mathcal{J}_1$:
  \begin{align*}
    \ADefine{
      \left(\HypJ{\mathcal{J}_2}{\mathcal{J}_1}\right)(\Upsilon)
    }{
      \left({\mathcal{J}_2}^{\mathcal{J}_1}\right)(\Upsilon)
    }\\
    &\cong
    \OMode{
      \Hom{\mathbb{J}}{
        \Yoneda{\Upsilon}\times\mathcal{J}_1
      }{
        \mathcal{J}_2
      }
    }\\
    &\cong
    \OMode{
      \int_{\Upsilon'}{{\Upsilon\inj\Upsilon'}
        \Rightarrow{{{\mathcal{J}}_{2}(\Upsilon')}^{{\mathcal{J}}_{1}(\Upsilon')}}}
    }
  \end{align*}

  In other words, the hypothetical judgment
  $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$ is evident when there shall forevermore
  be an effective transformation of evidence for the antecedent into evidence
  for the consequent. It is important to keep in mind that this is a
  \emph{semantic} consequence and therefore expresses admissibilities, in
  contrast to the \emph{logical} consequence which expresses
  derivabilities~\cite{harper:2016}. We will not have any need for the latter.

@@

  For any diagram $\Of{\mathcal{J}}{\mathbb{J}^{\square\Theta}}$, we can define
  the general judgment $\GenJ{\Theta}{\mathcal{J}}$ as follows, where
  $\Of{\Yoneda{-}}{\OpCat{\SCtx}\ra\mathbb{J}}$ is the Yoneda Embedding:
  \begin{align*}
    \ADefine{
      \left(\GenJ{\Theta}{\mathcal{J}}\right)(\Upsilon)
    }{
      \Hom{\mathbb{J}}{
        \Yoneda{\Upsilon}
      }{
        \varprojlim\mathcal{J}
      }
    }\\
    &\cong
    \OMode{
      \int_{(\Upsilon',\,\vec{E})}{\Upsilon\inj\Upsilon'\Rightarrow{\mathcal{J}(\vec{E})(\Upsilon')}}
    }
  \end{align*}

   Intuitively, the general judgment $\GenJ{\Theta}{\mathcal{J}}$ is evident
   when the family of judgments $\mathcal{J}$ shall forevermore have a global
   section (i.e.\ for any substitution $\Member{\vec{E}}{\square\Theta}$,
   $\mathcal{J}(\vec{E})$ shall be evident). Recall that metavariables have
   \emph{valences} rather than sorts; as such, the general judgment allows us
   to quantify over terms of higher type.

@@

\begin{definition}[Internal Judgments]

  It often happens that the evidence for a judgment is the knowledge of the
  meaning of a new form of judgment. This can be neatly expressed in our topos
  using the subobject classifier $\Omega$, where the evidence of $\Omega$ at
  $\Upsilon$ gives rise to an \emph{internal judgment} which is defined at all
  $\Upsilon'$ for which we have $\Of{\rho}{\Upsilon\inj\Upsilon'}$.

  A \emph{global internal judgment} is a section
  $\Of{J}{\int_\Upsilon{\Omega^{\mathbbold{1}}}(\Upsilon)}\cong
  \IMode{\int_\Upsilon\Hom{\mathbb{J}}{\Yoneda{\Upsilon}}{\Omega}}$, which can
  be externalized as a judgment $\Of{\Embed{J}}{\mathbb{J}}$ whose derivations
  at $\Upsilon$ are the arrows
  $\Of{\phi}{\Hom{\mathbb{J}}{\Yoneda{\Upsilon}}{\mathbbold{1}}}$ such that the
  following diagram commutes:

  \[
    \begin{tikzcd}
      \IMode{\Yoneda{\Upsilon}}
        \arrow[r, dashed, "\OMode{\phi}"]
        \arrow[rd, "\IMode{J_\Upsilon}"']
      & \IMode{\mathbbold{1}} \arrow[d, "\IMode{\mathbf{true}}"] \\
      & \IMode{\Omega}
    \end{tikzcd}
  \]

  In most cases, we will say something like ``To know $\mathcal{J}$ is to have
  a judgment $\mathcal{J'}$'' rather than appealing directly to the subobject
  classifier.
\end{definition}

@@

 \begin{remark}
   Formally, the evaluation relation is given in our logical framework by a
   global internal judgment $\Of{\Downarrow_\tau}{\int_\SCtx
   \Omega^{\Exprs_\tau\times\Values_\tau\times\NaturalsPsh}}$, where
   $\NaturalsPsh$ is the constant presheaf $\MapsTo{\Upsilon}{\Naturals}$.
   Then, we know $\EvalN{\Upsilon}{\tau}{n}{M}{N}$ just when the
   externalization $\Embed{\Downarrow_\tau}(M,N,n)$ is evident at $\Upsilon$.
 \end{remark}

>>>>

For any abt signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we can
deploy a generic logical framework equipped with higher-order judgments. A
judgment is defined by giving it a \emph{meaning explanation}, which is a
specification of the conditions under which it may be asserted (sc.\ its
evidence).

%% Recall from~\cite{sterling-morrison:2015} that
%% $\SCtx$ is the category of symbol-contexts and their injective, sort-preserving
%% renamings. Then, a \emph{judgment} shall be interpreted as an object in the presheaf
%% category $\Define{\mathbb{J}}{\Sets^\SCtx}$; in other words, a
%% judgment is identified with the intensional set of its renaming-sensitive
%% derivations. $\mathbb{J}$, like all presheaf categories, is a topos.

\begin{definition}[Hypothetical Judgment]
  For two judgments $\mathcal{J}_1$ and $\mathcal{J}_2$, we can form the
  hypothetical judgment $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$, whose evidence
  shall be an effective transformation of any evidence for $\mathcal{J}_1$ into
  evidence for $\mathcal{J}_2$. It is important to keep in mind that this is a
  \emph{semantic} consequence and therefore expresses admissibilities, in
  contrast to the \emph{logical} consequence which expresses
  derivabilities~\cite{harper:2016}. We will not have any need for the latter.

  %% using the
  %% exponential construction~\cite[p.~46]{maclane-moerdijk:1992}, which expresses
  %% the semantic consequence of $\mathcal{J}_2$ from $\mathcal{J}_1$:
  %% \begin{align*}
  %%   \ADefine{
  %%     \left(\HypJ{\mathcal{J}_2}{\mathcal{J}_1}\right)(\Upsilon)
  %%   }{
  %%     \left({\mathcal{J}_2}^{\mathcal{J}_1}\right)(\Upsilon)
  %%   }\\
  %%   &\cong
  %%   \OMode{
  %%     \Hom{\mathbb{J}}{
  %%       \Yoneda{\Upsilon}\times\mathcal{J}_1
  %%     }{
  %%       \mathcal{J}_2
  %%     }
  %%   }\\
  %%   &\cong
  %%   \OMode{
  %%     \int_{\Upsilon'}{{\Upsilon\inj\Upsilon'}
  %%       \Rightarrow{{{\mathcal{J}}_{2}(\Upsilon')}^{{\mathcal{J}}_{1}(\Upsilon')}}}
  %%   }
  %% \end{align*}

  %% In other words, the hypothetical judgment
  %% $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$ is evident when there shall forevermore
  %% be an effective transformation of evidence for the antecedent into evidence
  %% for the consequent.

@@

  For any $\square{\Theta}$-indexed family of judgments $\mathcal{J}$, we can
  express the general judgment $\GenJ{\Theta}{\mathcal{J}}$ as being evident
  when $\mathcal{J}(\vec{E})$ is evident for each
  $\Of{\vec{E}}{\square\Theta}$; in other words, the evidence of the general
  judgment is an effective transformation of $\Theta$-substitutions to their
  fibres in $\mathcal{J}$.  Recall that metavariables have \emph{valences}
  rather than sorts; as such, the general judgment allows us to quantify over
  terms of higher type.

%%
%%  For any diagram $\Of{\mathcal{J}}{\mathbb{J}^{\square\Theta}}$, we can define
%%  the general judgment $\GenJ{\Theta}{\mathcal{J}}$ as follows, where
%%  $\Of{\Yoneda{-}}{\OpCat{\SCtx}\ra\mathbb{J}}$ is the Yoneda Embedding:
%%  \begin{align*}
%%    \ADefine{
%%      \left(\GenJ{\Theta}{\mathcal{J}}\right)(\Upsilon)
%%    }{
%%      \Hom{\mathbb{J}}{
%%        \Yoneda{\Upsilon}
%%      }{
%%        \varprojlim\mathcal{J}
%%      }
%%    }\\
%%    &\cong
%%    \OMode{
%%      \int_{(\Upsilon',\,\vec{E})}{\Upsilon\inj\Upsilon'\Rightarrow{\mathcal{J}(\vec{E})(\Upsilon')}}
%%    }
%%  \end{align*}
%%
%%   Intuitively, the general judgment $\GenJ{\Theta}{\mathcal{J}}$ is evident
%%   when the family of judgments $\mathcal{J}$ shall forevermore have a global
%%   section (i.e.\ for any substitution $\Member{\vec{E}}{\square\Theta}$,
%%   $\mathcal{J}(\vec{E})$ shall be evident).

@@

%% \begin{definition}[Internal Judgments]
%%
%%   It often happens that the evidence for a judgment is the knowledge of the
%%   meaning of a new form of judgment. This can be neatly expressed in our topos
%%   using the subobject classifier $\Omega$, where the evidence of $\Omega$ at
%%   $\Upsilon$ gives rise to an \emph{internal judgment} which is defined at all
%%   $\Upsilon'$ for which we have $\Of{\rho}{\Upsilon\inj\Upsilon'}$.
%%
%%   A \emph{global internal (family of) judgments} is a section
%%   $\Of{J}{\int_\SCtx {\Omega^{\mathcal{J}}}}\cong
%%   \IMode{\int_\Upsilon\Hom{\mathbb{J}}{\Yoneda{\Upsilon}\times\mathcal{J}}{\Omega}}$,
%%   which can be externalized as a ``total'' judgment $\Of{\Embed{J}}{\mathbb{J}}$
%%   for whose derivations at $\Upsilon$ are the arrows
%%   $\Of{\phi}{\Hom{\mathbb{J}}{\Yoneda{\Upsilon}}{\mathbbold{1}\times\mathcal{J}}}$
%%   such that the following diagram commutes:
%%   \[
%%     \begin{tikzcd}[cramped,sep=huge]
%%       \IMode{\Yoneda{\Upsilon}}
%%         \arrow[r, dashed, "\OMode{\phi}"]
%%         \arrow[rd, "\IMode{J_\Upsilon\Pair{1}{\pi_2\circ\phi}}"']
%%       & \IMode{\mathbbold{1}\times\mathcal{J}} \arrow[d, "\IMode{\mathbf{true}\circ\pi_1}"] \\
%%       & \IMode{\Omega}
%%     \end{tikzcd}
%%   \]
%%
%%   Then, to say $\IMode{\Upsilon}\parallel\Embed{J}(E)$ where
%%   $\Member{E}{\mathcal{J}(\Upsilon)}$ is to assert that we have evidence
%%   $\Member{F}{\Embed{J}(\Upsilon)}$ such that $\IsEq{\pi_2(F)}{E}$.
%%
%%   When asserting the existence of an internal judgment, in most cases we will
%%   say something like ``To know $\mathcal{J}$ is to have a judgment
%%   $\mathcal{J'}$'' rather than appealing directly to the subobject classifier.
%% \end{definition}

@@
